### 2018-11-04 Sections/quantum_algorithms.tex
citations added minor fix

>>>>

\subsection{Quantum advantage with shallow circuits}\index{shallow_circuits}

As we discuss in Sec.~\ref{sec:NISQ}, we have good reasons to believe that quantum computer
have super-classical abilities. Recently, an unconditional proof of a quantum speed-up has
been shown for a class of circuits.

It is incredibly difficult to compare quantum VS classical polynomial time computation, and 
currently we have noisy gates. Given $n$ qubits, depth $d$, and an error rate $\epsilon$, intuitively, we need the condition
\begin{align}
nd \ll 1/\epsilon
\end{align}
\noindent to hold, in order for a quantum algorithm to run successfully. If $d$ is large, then noise dominates and our circuit is classically simulable; on the other hand, if we have a large $n$ and small $d$ (constant depth circuits), we may observe a quantum speed up in the near future. Given these factors, we are motivated to examine shallow circuits. 
Now we will see that there is a distinction between constant depth circuits run using classical VS quantum algorithm.

The result is for the following hidden linear function (HLF) problem \cite{bravyi2018quantum}. Given an 
$n\times n$ symmetric binary matrix $A$, we specify a quadratic form
\begin{align}
q(x) = \vec x^T A \vec x \text{(mod 4)}.
\end{align}
\noindent The goal is to find a binary vector $\vec z \in {0,1}^n$ such that
% 
\begin{align}
q(x) = 2 \vec z^T \vec x
\end{align}
\noindent for all $\vec x$ in the binary null-space of A.

The quantum alvorithm that solves the 2D HLF problem is similar to the Bernstein and Vazirani problem \cite{bernstein1997quantum}. Here a quantum circuit $U_q$ acts on the basis state $\vec x \in \{ 0,1\}^n$
\begin{align}
U_q \ket{x} = i^{q(x)} \ket{x}
\end{align}
The solution is obtained by measuring the state $\ket{\Psi_q}$ in the computational basis,
\begin{align}
\ket{\Psi_q} &= H^{\otimes n }U_q H^{\otimes n} \ket{0}^{\otimes n} \nonumber \\
&= \frac{1}{2^n}\sum_{\vec x,\vec z \in \{ 0,1\}^n} i^{q(x)} (-1)^{\vec z^T x}\ket{z}
\end{align}
\noindent where $H$ is the Hadamard gate. A schematic of the circuit is given in Fig~.\ref{fig:2DHLF}.
Destructive interference between terms in x results a vanishing amplitude
$\braket{z|\Psi_q}=0$ on basis state $\ket{z}$ which are not solutions of the HLF problem.
The circuit $U_q$ can be decomposed into a product of CZ gates and phase-shift gates 
\begin{align}
S_i = \begin{pmatrix}
    1 & 0\\
    0 & -i
    \end{pmatrix} \nonumber
\end{align}

\begin{figure}[h]
\includegraphics[trim = 0cm 0cm 0cm 0cm, clip, width=1.0\linewidth]{2DHLFcircuit.pdf}
\caption{\label{fig:2DHLF} Quantum circuit to solve the 2D HLF problem. }
\end{figure}

If we lay out the circuit in 2D (hence the 2D HLF problem), it only requires classically controlled Clifford gates between nearest neighbor qubits on a 2D grid. For an example, see Fig.~\ref{fig:2DHLFgrid}.
\begin{figure}[h]
\includegraphics[trim = 0cm 0cm 0cm 0cm, clip, width=0.8\linewidth]{2DHLF_grid.jpg}
\caption{\label{fig:2DHLFgrid} Implementation of the HLF problem on a 2D grid.}
\end{figure}

Now, comparing the quantum algorithm to a classical one, it has been proven that any probabilistic classical circuit with bounded input that solves the 2D HLF problem with high probability $\epsilon \leq 1/8$ must have depth increasing logarithmically with input size \cite{bravyi2018quantum}. The speed-up here is provided by quantum non-locality. 
% 
The fact that a problem can be solved with constant depth quantum circuits, but not classical ones shows that there is a clear separation in computational speed between the two cases.
